
\begin{acknowledgments}

First and foremost, I want to express my heartfelt gratitude to my advisor, Prof.\ Wen-mei Hwu. I was incredibly fortunate to receive an offer from him despite my initial challenges with English and inexperience. From the very beginning, he kept his door open. He incisively urged us to conduct profound research by identifying fundamental and scientific problems in real-world systems. He is a role model, demonstrating how to work as a scholar in all aspects. He showcases the invaluableness of commitment and perseverance. His unwavering support, wisdom, benevolence, and compassion have been constant sources of inspiration. Wen-mei generously provided research assistantship throughout our Ph.D.\ programs. He expanded our networks for collaboration and future careers. I am especially grateful that Wen-mei enabled us to explore potential problems freely and that he is patient with me even though I did not pursue many of the potential projects we found.

I am grateful to my final exam and prelim exam committees for their thoughtful feedback, which strengthened the dissertation.
I thank Prof.~Vikram Adve, Prof.\ Deming Chen, Prof.\ Steve Lumetta, and Prof.\ Sanjay Patel for serving on these committees.

Next, I would like to thank Dr.\ Dejan Milojicic, my internship manager and our collaborator. Like Prof.\ Izzat El Hajj, my internship at HP Labs was a turning point in my Ph.D.\ study. With Dejan's unyielding support and guidance, I practiced driving the research agenda, maintaining focus on key topics, connecting the dots, delivering convincing presentations, coordinating with other teams, etc. All these skills have been instrumental in conducting impactful research and pursuing a Ph.D.\ degree. In addition to his great empowerment, Dejan's insightful vision, whether shared in meetings or public presentations, has been a tremendous source of learning and inspiration.

Special thanks to Prof.\ Sitao Huang, Dr.\ Xiang Song, Dr.\ Xiaofan Zhang, Prof.\ Steve Lumetta, and Dr.\ Seung Won Min for their tremendous help during my search for dissertation topics.
In particular, Sitao led me into compiler research in the Pylog project. Through the PyTorch-Direct project led by Seung Won, I began optimizing the PyTorch stack for graph neural networks. This has been the starting point of our collaboration with the Amazon DGL team, which finally yielded the Hector work.
Without Sitao's and Seung Won's support, this dissertation would not have been possible. 
Xiang provided his utmost support during the ideation and execution of the Hector project, as did Xiaofan and Steve during the SSDTrain project. Their help was critical in laying the foundation for this dissertation.
Moreover, Steve, Xiaofan, and Sitao regularly provided constructive feedback on my dissertation writing, significantly elevating the quality of the final work. Additionally, I am grateful for help from Dr.\ Mert Hidayeto\u{g}lu, Dr.\ Zaid Qureshi, and Dr.\ Vikram Sharma Mailthody during my dissertation research.

Among the many invaluable opportunities that Wen-mei opens up and grants us, one is the chance to collaborate with many highly energetic and extraordinary scholars, including members of the Illinois Microarchitecture Project using Algorithms and Compiler Technology~(IMPACT) Group, professors at the University of Illinois, and industrial scientists.
I enjoyed and learned much from my close collaborators, e.g., Dr.\ Jeongmin Park, Dr.\ Da Zheng, Dr.\ Sai Rahul Chalamalasetti, and Dr.\ Israt Nisa.


I thank my software engineer intern hosts: Dr.\ Aart Bik, Dr.\ Penporn Koanantakool, Jerry Zheng, Dr.\ Howard Chen, James Player, Qingwei Lin, and Bo Qiao. The experience informed me of the focus of companies and, accordingly, how academic research could help or differentiate. Especially my time at Google helped me get started in large language models.

Before embarking on my Ph.D.\ journey, I was lucky to have been guided and mentored by many extraordinary scholars at Tsinghua and the University of California, Santa Barbara. Their insights and support equipped me to face the diverse challenges that arose during my dissertation research. I owe many thanks to Prof.\ Guohao Dai, Prof.\ Xing Hu, Dr.\ Shuangchen Li, Dr.\ Xinfeng Xie, Dr.\ Jie Fu, Prof.\ Yu Wang, Prof.\ Yuan Xie, and Prof.\ Guoqi Li.

Finally, I thank my girlfriend Peizhen Wu, who is also on her way to earning her Ph.D. Her enthusiastic support, belief in me, and knack for pacifying me when I am in trouble have been invaluable. More importantly, she suggests and exemplifies how to navigate the interpersonal aspects of research. 

\end{acknowledgments}
